Converging High-Throughput and High-Performance Computing: A Case Study

### 2017-04-02 nge.tex
edits, fixes and comments in 5.0 and 5.1

\jhanote{admittedly i haven't read earlier section in question  carefully, but
i'm a bit confused about the use of "wide area". wide area is typically used for
cross data center}\mtnote{Every PanDA Pilot has to communicate with PanDA
Server. PanDA Server runs on dedicated resources, outside OLCF. Titan's worker
nodes cannot communicate with PanDA Server because their connectivity is limited
to Titan's network. Do you have a suggestion for an alternative to ``wide
area''?} \jhanote{how is the following: "... deployed on the DTNs as Titan's worker
nodes lack connectivity to the wide area network"}.

As seen in \S\ref{sec:panda_deployment}, PanDA Broker was deployed on DTNs
because of the absence of wide area network connectivity on Titan's worker
nodes. The lack of pilot capabilities impacts both the efficiency and the
flexibility of PanDA's execution process. Pilots could help to improve
efficiency by increasing throughput and by enabling better use of backfill
utilization. Further, pilots makes it easier to support heterogeneous
workloads.

@@

The support of  multiple generations of workload would enable a more efficient
use of the backfill availability of walltime. Currently, when a set of
simulations ends, the PBS job also ends, independent of whether more wall-time
would still be available. With a pilot, additional simulations could be
executed  to utilize all the available wall-time, while avoiding further job
packaging and submission overheads.

@@

\jhanote{not sure MPI script is optimal description or ambigious.} Multiple
generations would also relax two assumptions of the current execution model:
knowing the number of simulations before submitting the MPI script, and having
a fixed number of events per simulation (100 at the moment). Pilots would
enable the scheduling of simulations independently from whether they were
available at the moment of submitting the pilot. Further, simulations with a
varying number of events could be scheduled on a pilot, depending on the
amount of remaining walltime and the distribution of execution time per event,
as shown in \S\ref{ssec:panda_titan}, Figure~\ref{fig:100event- distrib}.
These capabilities would increase the efficiency of the PanDA Broker when
there is a large difference between the number of cores and walltime.

@@

NGE is a runtime system designed to execute heterogeneous and dynamic workloads
on diverse resources. Fig.~\ref{fig:arch-overview} illustrates its current
architecture as deployed on Titan: the two management modules represent a
simplified version of the PanDA Broker while the agent module is the pilot
submitted to Titan and executed on its worker nodes. The communication between
PanDA Broker and PanDA Server is abstracted away as it is not immediately useful to
evaluate the performance and capabilities of a pilot on Titan.

@@

NGE exposes an API for the application layer to describe workloads
(Fig.~\ref{fig:arch-overview}, green squares) and pilots
(Fig.~\ref{fig:arch-overview}, red circles), and to instantiate a PilotManager
and a UnitManager. The PilotManager submits pilots to Titan's PBS batch system
via SAGA API (Fig.~\ref{fig:arch-overview}, dash arrow), as done by PanDA Broker
to submit MPI scripts. Once scheduled, the pilot's Agent is bootstrapped on
Titan's MOM node and worker nodes, and  the UnitManager schedules units to the
Agent's Scheduler (Fig.~\ref{fig:arch-overview}, solid arrow). The Agent's
Scheduler schedules the units on one or more Agent's Executor for execution. The
Agent's executors can manage one or more worker nodes, depending on performance
evaluations.

The UnitManager and the Agent communicate via a database that is instantiated on
one of Titan's DTN so as to be reachable by both modules. A similar approach would
be used to enable PanDA Broker to schedule pilots instead of MPI
scripts.\mtnote{Rewrite this after checking it with Andre.}

@@

The NGE Agent uses the \emph{Open Run-Time Environment (ORTE)} for communication
and coordination of units execution. ORTE is a spin-off from the Open-MPI
project and is a critical component of the OpenMPI implementation. It was
developed to support distributed high-performance computing applications
operating in a heterogeneous environment. The system transparently provides
support for interprocess communication, resource discovery and allocation, and
process launching across a variety of platforms. ORTE provides a mechanism
similar to the Pilot concept - it allows the user to create a \emph{``dynamic
virtual machine''} (DVM) that spans multiple nodes. ORTE provides libraries to
enable the submission, monitoring and managing of tasks, avoiding filesystem
bottlenecks and race conditions with network sockets. As a consequence, it is
able to minimize the system overhead while submitting tasks.

>>>>

% \jhanote{admittedly i haven't read earlier section in question  carefully, but
% i'm a bit confused about the use of "wide area". wide area is typically used for
% cross data center}\mtnote{Every PanDA Pilot has to communicate with PanDA
% Server. PanDA Server runs on dedicated resources, outside OLCF. Titan's worker
% nodes cannot communicate with PanDA Server because their connectivity is limited
% to Titan's network. Do you have a suggestion for an alternative to ``wide
% area''?} \jhanote{how is the following: "... deployed on the DTNs as Titan's worker
% nodes lack connectivity to the wide area network"}.

As seen in \S\ref{sec:panda_deployment}, PanDA Broker was  deployed on the
DTNs as Titan's worker nodes lack connectivity to the wide area network The
lack of pilot capabilities impacts both the efficiency and the flexibility of
PanDA's execution process. Pilots could improve efficiency by increasing
throughput and enabling greater backfill utilization. Further, pilots makes it
easier to support heterogeneous workloads.

@@

The support of  multiple generations of workload would enable more efficient
use of the backfill availability of walltime. Currently, when a set of
simulations ends, the PBS job also ends, independent of whether more wall-time
would still be available. With a pilot, additional simulations could be
executed  to utilize all the available wall-time, while avoiding further job
packaging and submission overheads.

@@

\jhanote{not sure MPI script is optimal description or ambigious.} Multiple
generations would also relax two assumptions of the current execution model:
knowing the number of simulations before submitting the MPI script, and having
a fixed number of events per simulation (100 at the moment). Pilots would
enable the scheduling of simulations independently from whether they were
available at the moment of submitting the pilot. Further, simulations with a
varying number of events could be scheduled on a pilot, depending on the
amount of remaining walltime and the distribution of execution time per event,
as shown in \S\ref{ssec:panda_titan}, Figure~\ref{fig:100event-distrib}.
These capabilities would increase the efficiency of the PanDA Broker when
there is a large difference between the number of cores and walltime.

@@

\jhanote{Alternative: "NGE is a runtime system to execute heterogeneous and
dynamically determined tasks that constitute workloads." This definition put
the focus on tasks and not workloads, which is a level above in the RADICAL
conceptual layer diagram.} NGE is a runtime system designed to execute
heterogeneous and dynamic workloads on diverse resources. Fig.~\ref{fig:arch-
overview} illustrates its current architecture as deployed on Titan: the two
management modules (Pilot and Unit) represent a simplified version of the PanDA Broker while
the agent module is the pilot submitted to Titan and executed on its worker
nodes. The communication between PanDA Broker and PanDA Server is abstracted
away as it is not immediately useful to evaluate the performance and
capabilities of a pilot on Titan.

@@

NGE exposes an API for the application  \jhanote{workload management?} layer
to describe workloads (Fig.~\ref{fig:arch-overview}, green squares) and pilots
(Fig.~\ref{fig:arch-overview}, red circles), and to instantiate a PilotManager
and a UnitManager.   The PilotManager submits pilots to Titan's PBS
\jhanote{for consistency should we say Moab?} batch system via SAGA API
(Fig.~\ref{fig:arch-overview}, dash arrow), as done by PanDA Broker to submit
MPI scripts. Once scheduled, the pilot's Agent is bootstrapped on Titan's MOM
node and \jhanote{the Agent's Executors on the} worker nodes, and  the
UnitManager schedules units to the Agent's Scheduler (Fig.~\ref{fig:arch-
overview}, solid arrow). The Agent's Scheduler schedules the units on one or
more Agent's Executor for execution. The Agent's executors can manage one or
more worker nodes, depending on performance evaluations.

The UnitManager and the Agent communicate via a database that is instantiated on
one of Titan's DTN so as to be reachable by both modules. A similar approach would
be used to enable PanDA Broker to schedule pilots instead of MPI
scripts.\mtnote{Rewrite this after checking it with Andre.}\jhanote{In the interest of space and clarity, I would  remove the last sentence of this para. And merge
the first sentence with the para above.}

@@

The NGE Agent uses the \emph{Open Run-Time Environment (ORTE)} for
communication and coordination of the execution of units. ORTE is a critical
component of the OpenMPI implementation, developed to support distributed
high-performance computing applications operating in a heterogeneous
environment~\cite{XXX}. The system \jhanote{which system? ORTE or OpenMPI in
general?} transparently provides support for interprocess communication,
resource discovery and allocation, and process launching across a variety of
platforms. ORTE provides a mechanism similar to the Pilot concept - it allows
the user to create a \emph{``dynamic virtual machine''} (DVM) that spans
multiple nodes. ORTE provides libraries to enable the submission, monitoring
and managing of tasks, avoiding filesystem bottlenecks and race conditions
with network sockets. As a consequence, it is able to minimize the system
overhead while submitting tasks.
